manuscript_figures

### 2024-06-28 scripts/manuscript_figures.Rmd
calculated PERMANOVA and SIMPER results

@@

>>>>

# Save the plot to a file
ggsave("output/figures/nmds.plot.png", plot = nmds_plot, width = 15, height = 10)

@@

PERMANOVA and SIMPER analysis
```{r}
# Remove rows where all abundance values are zero
non_zero_rows <- rowSums(community_data) > 0
community_data <- community_data[non_zero_rows, ]
environmental_data <- environmental_data[non_zero_rows, ]

# Convert YEAR to factor
environmental_data$YEAR <- as.factor(environmental_data$YEAR)

# Perform PERMANOVA
permanova_result <- adonis2(community_data ~ SITE + YEAR, data = environmental_data, permutations = 999, method = "bray")

# View the results
print(permanova_result)

# For the SIMPER analysis, we'll use the original (non-transformed) data, but still excluding the all-zero rows
simper_result <- simper(community_data, group = environmental_data$SITE, permutations = 999)

# View summary of results
summary(simper_result)

# Extract the relevant information from the SIMPER results
simper_data <- data.frame(
  Species = simper_result$`Fort Wetherill_King's Beach`$species,
  Contribution = simper_result$`Fort Wetherill_King's Beach`$average,
  Fort_Wetherill = simper_result$`Fort Wetherill_King's Beach`$ava,
  Kings_Beach = simper_result$`Fort Wetherill_King's Beach`$avb
)

# Calculate the percentage contribution
simper_data$Percentage <- simper_data$Contribution / sum(simper_data$Contribution) * 100

# Reshape the data for plotting
simper_long <- tidyr::pivot_longer(simper_data, 
                                   cols = c(Fort_Wetherill, Kings_Beach),
                                   names_to = "Site", 
                                   values_to = "Abundance")

# Create the plot
ggplot(simper_long, aes(x = Species, y = Abundance, fill = Site)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_text(data = simper_data, 
            aes(x = Species, y = max(simper_long$Abundance) + 0.5, 
                label = sprintf("%.1f%%", Percentage)),
            position = position_dodge(width = 0.9),
            vjust = -0.5) +
  labs(title = "Species Contributions to Differences Between Sites",
       subtitle = "Based on SIMPER analysis",
       y = "Average Abundance",
       x = "Species") +
  scale_fill_manual(values = c("Fort_Wetherill" = "grey", "Kings_Beach" = "black")) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.title = element_blank())

PERMANOVA results:
adonis2(formula = community_data ~ SITE + YEAR, data = environmental_data, permutations = 999, method = "bray")
         Df SumOfSqs      R2      F Pr(>F)    
SITE      1   0.2632 0.02501 1.3160  0.235    
YEAR      4   2.8619 0.27191 3.5774  0.001 ***
Residual 37   7.4000 0.70308                  
Total    42  10.5251 1.00000

SIMPER results:
Contrast: Fort Wetherill_King's Beach 
            average      sd   ratio     ava     avb cumsum     p  
molluscs    0.56710 0.30146 1.88120 7.64700 5.84300  0.894 0.803  
echinoderms 0.04820 0.10148 0.47480 0.14700 0.34300  0.970 0.067 .
crustacean  0.01920 0.04652 0.41350 0.30000 0.03700  1.000 0.713  
---
Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1
Permutation: free
Number of permutations: 999
